---
title: "ctafinal"
output:
  word_document: default
  html_document: default
date: "2026-04-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Load packages
library(tidyverse) #Tidy-style text mining; provides unnest_tokens() for tokenization and bind_tf_idf() for TF-IDF.Used in Step 1 (tokenizing speeches) and Step 2 (computing inter-party cosine similarity).

library(lubridate) # Simplifies parsing and manipulating date/time objects

library(lme4) #Fits linear and generalized linear mixed-effects models.
library(lmerTest)   
library(broom.mixed) # Tidies output from mixed-effects models (lme4/lmerTest) into data frames via tidy().
library(dplyr)
library(tidytext) 
library(ggplot2) #visualization engine
library(patchwork) #Combine multiple ggplot2 plots into a single layout
library(conText)
library(quanteda)
library(dplyr)
library(tidyr)
library(readr)
library(stringr)
library(quanteda.textstats)

```

```{r}
#preparation 
#read the dataset and exlore the data struction
uk_data <- read_csv("uk_data.csv")
glimpse(uk_data)
```

```{r}
# filter data 
# select data from 2010 to 2019, exclude 2016 to prepare for the comparison with Pre-Brexit and Post- Brexit. also, the speeches should belong to either Labour or Conservative.
uk_analysis_data <- uk_data %>%
  mutate(year = year(date)) %>%
  filter(party %in% c("Labour", "Conservative")) %>%
  filter(year >= 2010 & year <= 2019 & year != 2016) %>%
  # name the period of Pre-Brexit as Pre, Post-Brexit as Post.
  mutate(period = case_when(
    year >= 2010 & year <= 2015 ~ "Pre",
    year >= 2017 & year <= 2019 ~ "Post"
  ))

# check the count of the speech
uk_analysis_data %>%
  count(party, period)
```
```{r}
# # get the Extended Moral Foundations Dictionary (eMFD).(https://github.com/medianeuroscience/emfd)
emfd_dict <- read_csv("emfd_scoring.csv")
head(emfd_dict)
```


```{r}
# select necessary varaiables
uk_clean <- uk_analysis_data %>%
  select(id_speech, date, period, party, text, top_topic, government, last_name, words)
```

```{r}
#manipulate the data into tidy format again, unnesting each token (here: words) from speech text.
uk_tokens <- uk_clean %>%
  unnest_tokens(word, text) %>%
  group_by(id_speech) %>%
  ungroup()
```

```{r}
# match eMFD（stop words and digits are excluded naturally）
uk_scored_words <- uk_tokens %>%
  inner_join(emfd_dict, by = "word")
```

```{r}
#Step1. eMFD Moral Foundation Scoring
# Speech-level aggregation: moral foundation density (per 100 words).Determine which top topics exhibit the highest moral density.
uk_moral_density <- uk_scored_words %>%
  group_by(id_speech, party, period, top_topic, words) %>%
  summarise(
    care      = sum(care_p)      / first(words) * 100,
    fairness  = sum(fairness_p)  / first(words) * 100,
    loyalty   = sum(loyalty_p)   / first(words) * 100,
    authority = sum(authority_p) / first(words) * 100,
    sanctity  = sum(sanctity_p)  / first(words) * 100,
    .groups = "drop"
  )

```

```{r}
# Party × period summary for descriptive overview
density_comparison <- uk_moral_density %>%
  group_by(party, period) %>%
  summarise(across(care:sanctity, mean, na.rm = TRUE),
            .groups = "drop")

print(density_comparison)
```

```{r}
#visualize the outcome
# Four plots:
#     fig1. Yearly trend lines showing how each foundation's density evolves
#         for Labour vs Conservative, with Brexit 2016 as reference line.
#     fig2. A topic-level heatmap showing which debate topics carry the
#         highest moral density for each party.
#     fig3. Change in Moral Foundation Density: Post vs. Pre-Brexit.
#     fig4. DiD coefficients (is_con × post) 

# (a) Yearly trend lines
#merge the moral foundation scores with metadata (year) using a left join to retain all observations from the main dataset.
yearly_moral <- uk_moral_density %>%
  left_join(
    uk_analysis_data %>% select(id_speech, year),
    by = "id_speech"
  ) %>%
  #aggregate the data by party and year, calculating the mean score for each moral foundation dimension
  group_by(party, year) %>%
  summarise(across(care:sanctity, mean, na.rm = TRUE), .groups = "drop") %>%
  pivot_longer(care:sanctity, names_to = "foundation", values_to = "score")

p_trend <- ggplot(yearly_moral, aes(x = year, y = score, color = party)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.8) +
  #A vertical reference line is added at 2016 to indicate the Brexit referendum as a key political event.
  geom_vline(xintercept = 2016, linetype = "dashed", color = "grey50") +
  annotate("text", x = 2016.2, y = Inf,
           label = "Brexit\n2016", vjust = 1.5, hjust = 0,
           size = 3, color = "grey40") +
  #Each subplot represents a different moral foundation dimension.Free y-scales are used to allow variation across foundations with different value ranges.
  facet_wrap(~ foundation, scales = "free_y", ncol = 2) +
  scale_x_continuous(breaks = seq(2010, 2019, 2)) +
  scale_color_manual(values = c("Labour" = "#F4D35E", "Conservative" = "#4A90E2")) +
  labs(title = "Moral Foundation Density by Party (2010–2019)",
       subtitle = "Mean score per 100 words; 2016 excluded from data (referendum year)",
       x = NULL, y = "Score (per 100 words)", color = NULL) +
  labs(caption = "Fig. 1. Moral foundation density by party (2010–2019)") +
  theme_minimal() +
  theme(legend.position = "bottom",
        strip.text = element_text(face = "bold")) +
  theme(panel.spacing = unit(2, "lines")
)
p_trend

```
#explanation of Fig1.
# Figure 1 plots each foundation's density year-by-year with a dashed vertical line at 2016. The trend lines tell a consistent story: on most foundations, the Conservative line sits above or converges with the Labour line before 2016 and clearly separates upward from it afterward, most visibly on loyalty and authority, where the gap widens from roughly 0.05 to over 0.15 points per 100 words by 2018–2019. A flat Labour trajectory against a rising Conservative trajectory is exactly the visual signature a DiD design is designed to exploit. The figure also provides informal evidence for the parallel-trends assumption that DiD requires: in the Pre period (2010–2015), the two parties move in parallel on all five foundations, with no systematic divergence building up before the treatment. 


```{r}
# (b) Topic × foundation heatmap
topic_moral <- uk_moral_density %>%
  filter(!is.na(top_topic), top_topic != "no topic") %>%
  #The data is grouped by topic and party, and the mean score for each moral foundation dimension is calculated. The number of observations per group (n) is also retained.
  group_by(top_topic, party) %>%
  summarise(across(care:sanctity, mean, na.rm = TRUE),
            n = n(), .groups = "drop") %>%
  
  #Groups with fewer than 50 observations are excluded to reduce noise and improve the reliability of the estimates
  filter(n > 50) %>%
  pivot_longer(care:sanctity, names_to = "foundation", values_to = "score")

p_heatmap <- ggplot(topic_moral,
                    aes(x = foundation,
                        y = reorder(top_topic, score),#Topics are reordered based on their scores to improve interpretability and visual clarity in the heatmap
                        fill = score)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(score, 2)), size = 2.3) +
  scale_fill_gradient(low = "#E3F2FD", high = "#1565C0") +
  #Faceting by party allows comparison of moral expression patterns across political groups.
  facet_wrap(~ party) + #
  labs(title = "Moral Density by Topic and Party",
       x = NULL, y = NULL, fill = "Score") +
  labs(caption = "Fig. 2. Moral Density by Topic and Party ") +
  theme_minimal(base_size = 10) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
p_heatmap
```
#explanation of Fig2.
#Figure 2 disaggregates moral density by the 20-topic top-topic label rather than by year. It serves two purposes. First, it confirms that moral density is concentrated in topics where one would expect moral vocabulary "fabric of society", "law and order", "social services"  rather than distributed evenly, which is a basic face-validity check on the eMFD. Second, it shows that Conservative and Labour draw on similar moral repertoires on the same topic (the row-wise patterns across the two facets are highly correlated), but the intensity differs systematically: Conservative cells are on average 0.05–0.15 points darker than the corresponding Labour cell on almost every (topic × foundation) combination. Moral topical specialisation by party therefore appears weak; moral intensity by party appears to differ. This already narrows the answer to the RQ: polarisation is more likely a question of quantity than of distinct moral agendas.

```{r}
# Pre→Post Change Heatmap by Topic and Party 
# Computes the change in moral foundation density from Pre (2010–2015) to
#     Post (2017–2019) for each topic × party combination. Blue = increased
#     post-Brexit, Red = decreased. This reveals WHICH topics and foundations
#     shifted most, and whether the two parties shifted in opposite directions

# 1.Aggregate by topic × party × period; filter low-count cells
topic_period_moral <- uk_moral_density %>%
  filter(!is.na(top_topic), top_topic != "no topic") %>%
  group_by(top_topic, party, period) %>%
  summarise(across(care:sanctity, mean, na.rm = TRUE),
            n = n(), .groups = "drop") %>%
  filter(n > 30)   
# Threshold lowered to 30 because splitting by period reduces cell sizes.

#  2: Pivot to long then wide to compute Post - Pre difference
topic_diff <- topic_period_moral %>%
  pivot_longer(care:sanctity,
               names_to  = "foundation",
               values_to = "score") %>%
  pivot_wider(
    id_cols    = c(top_topic, party, foundation),
    names_from = period,
    values_from = score
  ) %>%
  mutate(diff = Post - Pre) %>%
  filter(!is.na(diff))
# Rows where either Pre or Post is missing (due to low n) are dropped.


# 3: Plot
ggplot(topic_diff,
       aes(x    = foundation,
           y    = reorder(top_topic, diff),
           fill = diff)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = round(diff, 2)), size = 2.5) +
  #Diverging colour scale: blue = Post increase, red = Post decrease.。
  scale_fill_gradient2(low     = "#CC0000",
                       mid     = "white",
                       high    = "#003F87",
                       midpoint = 0) +
  #  directly compare Conservative vs Labour shifts side by side for the same topic-foundation cell.
  facet_wrap(~ party) +
  scale_x_discrete(labels = c("authority" = "authority",
                               "care"      = "care",
                               "fairness"  = "fairness",
                               "loyalty"   = "loyalty",
                               "sanctity"  = "sanctity")) +
  labs(
    title    = "Change in Moral Foundation Density: Post vs. Pre-Brexit",
    subtitle = "Blue = increased post-Brexit; Red = decreased (Post minus Pre, per 100 words)",
    caption  = "Fig. 3. Pre→Post change by topic and moral foundation, split by party.",
    x        = NULL,
    y        = NULL,
    fill     = "Δ Score"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x  = element_text(angle = 30, hjust = 1),
    strip.text   = element_text(face = "bold"),
    legend.position = "right"
  )
```

#explanation of Fig3.
#Figure 3 isolates the Post-minus-Pre change in moral density within each (topic × foundation × party) cell. The colour asymmetry between the two facets is the key pattern: on topics such as "fabric of society", "law and order", and "finance and economy", the Conservative facet shows substantially more dark-blue (increase) cells particularly on loyalty, fairness and authority , whereas the Labour facet is paler and shows a wider mix of small increases and small decreases. The fact that the two parties shifted in different directions on the same topic–foundation cells (e.g. increased loyalty on "law and order" for Conservatives but near-zero change for Labour) is the visual form of a DiD interaction effect, which I now test formally.

```{r}
# Step 1 Statistical Test — Difference-in-Differences (DiD) 
# Tests whether Conservative and Labour diverged in moral foundation usage after Brexit, controlling for MP-level and topic-level variance.
#     The key coefficient is is_con:post — the DiD interaction term.
#     Positive = Conservative increased MORE (or decreased LESS) than Labour
#     post-Brexit on that foundation. This is the statistical confirmation of
#     the quantity divergence suggested by the trend lines (Fig. 1).

# Merge moral density scores with covariates needed for regression
reg_data <- uk_moral_density %>%
  left_join(
    uk_analysis_data %>%
      select(id_speech, year, id_mp, government, pm_questions, female, age),
    by = "id_speech"
  ) %>%
  mutate(
    # Binary indicator , Conservate=1, Labour=0
    is_con     = as.integer(party == "Conservative"),
    # Binary indicator , Post-Brexit = 1，Pre-Brexit = 0
    post       = as.integer(period == "Post"),
    #  DiD interaction term — the key variable of interest
    is_con_post = is_con * post,
    id_mp      = as.character(id_mp)
  )

```


```{r}
#  Run mixed-effects DiD for each moral foundation separately
#  Random intercepts for MP (id_mp) control for individual speaking styles;
#     random intercepts for topic (top_topic) control for topic-level moral framing.

foundations <- c("care", "fairness", "loyalty", "authority", "sanctity")

did_results <- map_dfr(foundations, function(f) {
  formula <- as.formula(paste0(
    f, " ~ is_con + post + is_con_post + ",
     "pm_questions + (1 | id_mp) + (1 | top_topic)"
  ))
  m <- lmer(formula, data = reg_data, REML = TRUE)
  tidy(m, effects = "fixed") %>%
    filter(term == "is_con_post") %>%
    mutate(foundation = f)
})

# Print coefficient table
did_results %>%
  select(foundation, estimate, std.error, statistic, p.value) %>%
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE            ~ "n.s."
    )
  ) %>%
  print()

```

```{r}
#  Forest plot visualisation (Fig. 4)
#  Each point is the DiD coefficient (is_con:post) for one moral foundation,
#     with 95% confidence intervals. Points to the right of zero mean Conservative
#     increased more than Labour post-Brexit on that foundation.

ggplot(did_results,
       aes(x     = reorder(foundation, estimate),
           y     = estimate,
           color = p.value < 0.05)) +
  geom_point(size = 3.5) +
  geom_errorbar(aes(ymin = estimate - 1.96 * std.error,
                    ymax = estimate + 1.96 * std.error),
                width = 0.2, linewidth = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  coord_flip() +
  scale_color_manual(
    values = c("TRUE"  = "#003F87",
               "FALSE" = "grey60"),
    labels = c("TRUE"  = "p < 0.05",
               "FALSE" = "n.s.")
  ) +
  labs(
    title    = "Difference-in-Differences: Party × Post-Brexit",
    subtitle = "Positive = Conservative used more relative to Labour post-Brexit",
    caption  = "Fig. 4. DiD coefficients (is_con × post) with 95% CIs.\nControls: government role, PM questions; random intercepts for MP and topic.",
    x        = NULL,
    y        = "DiD Coefficient (Conservative × Post)",
    color    = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    plot.caption    = element_text(hjust = 0, color = "grey40")
  )
```

#explanation of Fig4.
#Figure 4 presents the DiD coefficients as a forest plot with 95% confidence intervals. All five foundations show positive coefficients, with loyalty and fairness exhibiting the largest effects. The authority coefficient is the smallest and its confidence interval nearly touches zero.
```{r}
# Step2 Semantic Similarity
# adapts the TF-IDF cosine-similarity framework originally developed by Liang (2014, p. 160) to quantify "shared knowledge" in political discussion.
# Computes inter-party cosine similarity of TF-IDF vectors for each year in two versions:
#     (A) Full vocabulary , to capture overall discourse distance between parties
#     (B) Moral vocabulary only (eMFD words) , to capture distance specifically in the moral domain
#     then comparing the two curves to argue that Brexit polarised moral discourse specifically, not just general parliamentary language
#     Note: 2016 is retained here for visual continuity only — it is not used in any Pre/Post grouping.
# Prepare: collapse all speeches into one text per party per year
party_year_text <- uk_data %>%
  mutate(year = year(date)) %>%
  filter(
    party %in% c("Labour", "Conservative"),
    year >= 2010 & year <= 2019
  ) %>%
  group_by(party, year) %>%
  summarise(combined_text = paste(text, collapse = " "),
            .groups = "drop")
```

```{r}
# (A) full vocabulary
tfidf_full <- party_year_text %>%
  unnest_tokens(word, combined_text) %>%
  filter(!str_detect(word, "^[0-9]+$")) %>%
  anti_join(       
    # stop_words's choice (snowball) is following Gentzkow et al.(2019.p.7)'s research about linguistic differences between political groups 
    stop_words %>% filter(lexicon == "snowball"),
    by = "word"
  ) %>%
  mutate(doc_id = paste(party, year, sep = "_")) %>%
  count(doc_id, party, year, word, name = "n") %>%
  #Compute TF-IDF weights for each word in each document.
  bind_tf_idf(word, doc_id, n)

```

```{r}
#Keep only the variables needed for similarity calculation
cosine_full <- tfidf_full %>%
  select(party, year, word, tf_idf) %>%
#Reshape data so that each word has TF-IDF values for both parties.
  pivot_wider(names_from  = party,
              values_from = tf_idf,
              values_fill = 0) %>%
  group_by(year) %>%
#Compute cosine similarity between Labour and Conservative TF-IDF vectors.The format is following Liang (2014, p.160)'s SS(u1, u2) = <phi(u1), phi(u2)> / (||phi(u1)|| * ||phi(u2)||)
  summarise(
    cosine_sim = sum(Labour * Conservative) /
      (sqrt(sum(Labour^2)) * sqrt(sum(Conservative^2))),
    .groups = "drop" ) %>%
  mutate(version = "Full vocabulary")
```

```{r}
# (B) Moral-vocabulary-only cosine similarity

# Restrict to eMFD words only before computing TF-IDF. This ensures the cosine similarity reflects only the moral domain of discourse, directly operationalising the RQ's focus on moral vocabulary.

tfidf_moral <- party_year_text %>%
  unnest_tokens(word, combined_text) %>%
  inner_join(emfd_dict %>% select(word), by = "word") %>%
  # inner_join keeps only eMFD words ,so is no need to remove other stop words
  mutate(doc_id = paste(party, year, sep = "_")) %>%
  count(doc_id, party, year, word, name = "n") %>%
  bind_tf_idf(word, doc_id, n)
# the following steps are similar to the full vocabulary version above
cosine_moral <- tfidf_moral %>%
  select(party, year, word, tf_idf) %>%
  pivot_wider(names_from  = party,
              values_from = tf_idf,
              values_fill = 0) %>%
  group_by(year) %>%
  summarise(
    cosine_sim = sum(Labour * Conservative) /
      (sqrt(sum(Labour^2)) * sqrt(sum(Conservative^2))),
    .groups = "drop"
  ) %>%
  mutate(version = "Moral vocabulary (eMFD)")

```

```{r}
# Combine and plot (Fig. 5 Semantic Similarity)

cosine_combined <- bind_rows(cosine_full, cosine_moral)

ggplot(cosine_combined,
       aes(x = year, y = cosine_sim,
           color = version, linetype = version)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_vline(xintercept = 2016,
             linetype = "dashed", color = "grey50") +
  annotate("text", x = 2016.15, y = Inf,
           label = "Brexit\n2016", vjust = 1.5, hjust = 0,
           size = 3, color = "grey40") +
  annotate("rect", xmin = 2010, xmax = 2015,
           ymin = -Inf, ymax = Inf,
           alpha = 0.04, fill = "#003F87") +
  annotate("rect", xmin = 2017, xmax = 2019,
           ymin = -Inf, ymax = Inf,
           alpha = 0.04, fill = "#CC0000") +
  scale_x_continuous(breaks = 2010:2019) +
  scale_color_manual(values = c(
    "Full vocabulary"       = "grey40",
    "Moral vocabulary (eMFD)" = "#003F87"
  )) +
  scale_linetype_manual(values = c(
    "Full vocabulary"       = "dotted",
    "Moral vocabulary (eMFD)" = "solid"
  )) +
  labs(
    title    = "Inter-Party Semantic Similarity: Full vs. Moral Vocabulary",
    subtitle = "Cosine similarity of TF-IDF vectors (Labour vs. Conservative), 2010–2019",
    caption  = "Fig. 5. Blue shading = Pre-Brexit (2010–2015); Red shading = Post-Brexit (2017–2019).\nSolid line = eMFD moral words only; Dotted line = full vocabulary.",
    x        = NULL,
    y        = "Cosine Similarity",
    color    = NULL,
    linetype = NULL
  ) +
  theme_minimal() +
  theme(
    legend.position    = "bottom",
    panel.grid.minor   = element_blank(),
    plot.caption       = element_text(hjust = 0, color = "grey40")
  )
```

# explanation of fig5
#  Figure 5 plots two cosine similarity curves. The full-vocabulary curve rises from 0.37 in 2010 to over 0.93 by 2019, which likely reflects a shared agenda effect: after the referendum both parties debated the same Brexit-related legislation using much of the same technical and procedural language. The moral-vocabulary curve tells a different story. It drops from 0.85 in 2015 to 0.51 in 2017, the lowest point in the series. 

# This is the key cross-validation of the Step 1 result. If the DiD effect in Fig 4. simply reflected a general redrawing of parliamentary language post-Brexit, Step 2's two lines would move together. The fact that the moral-only line diverges faster than the full line is consistent with a moral polarisation claim, not a generic-polarisation claim.


```{r}
# Step 3 Semantic Divergence
# Select target words for ALC analysis
# 3.1 for every eMFD word, assign its dominant foundation = the foundation with the highest probability among the five dimensions
emfd_dominant <- emfd_dict %>%
  mutate(dominant = case_when(
    loyalty_p >= care_p & loyalty_p >= fairness_p & 
      loyalty_p >= authority_p & loyalty_p >= sanctity_p ~ "loyalty",
    fairness_p >= care_p & fairness_p >= loyalty_p & 
      fairness_p >= authority_p & fairness_p >= sanctity_p ~ "fairness",
    care_p >= fairness_p & care_p >= loyalty_p & 
      care_p >= authority_p & care_p >= sanctity_p ~ "care",
    authority_p >= care_p & authority_p >= fairness_p & 
      authority_p >= loyalty_p & authority_p >= sanctity_p ~ "authority",
    TRUE ~ "sanctity"
  ))

```

```{r}
# pick loyalty and fairness.Because among the five moral foundations (see the DID effect from step1), loyalty and fairness exhibit the largest and most statistically significant interaction effects (loyalty: β = 0.078, p < 0.001; fairness: β = 0.074, p < 0.001), indicating that these two dimensions show the strongest differential increase in moral language density between Conservative and Labour after Brexit. While sanctity also reaches statistical significance (β = 0.039, p < 0.001), its effect size is roughly half that of loyalty and fairness, suggesting a more modest partisan divergence.
loyalty_words <- emfd_dominant %>% filter(dominant == "loyalty") %>% pull(word)
fairness_words <- emfd_dominant %>% filter(dominant == "fairness") %>% pull(word)

# Filter the eMFD dictionary to keep only words whose dominant foundation is loyalty or fairness
lf_words_strict <- emfd_dominant %>%
  filter(dominant %in% c("loyalty", "fairness")) %>%
# Key filter: the word's probability on its dominant dimension must exceed 0.3 (my own choice)
# This removes words like "minister" where loyalty is technically dominant but loyalty_p is only 0.04
  filter(
    (dominant == "loyalty" & loyalty_p > 0.3) |
    (dominant == "fairness" & fairness_p > 0.3)
  ) %>%
  select(word, dominant)
```

```{r}
# Within the surviving set, pick the top-5 words per foundation by the minmum frequency across the two parties.
#   Taking the minimum ensures the word has enough data on BOTH sides for stable group-level embeddings.

top_words_strict <- uk_scored_words %>%
  # Keep only words that passed the strict moral threshold
  filter(word %in% lf_words_strict$word) %>%
  filter(party %in% c("Conservative", "Labour")) %>%
  # Count occurrences of each word per party
  count(party, word) %>%
   # Reshape: one row per word, one column per party
  pivot_wider(names_from = party, values_from = n, values_fill = 0) %>%
  # Take the minimum of the two parties' counts
  mutate(min_freq = pmin(Conservative, Labour)) %>%
  left_join(lf_words_strict, by = "word") %>%
  # Group by foundation, pick top 5 by min_freq within each
  group_by(dominant) %>%
  slice_max(min_freq, n = 5) %>%
  arrange(dominant, desc(min_freq))

print(top_words_strict)

```

```{r}
#3.2 GloVe: Stanford GloVe 6B 300-dimensional word embeddings (400k vocab),
# pre-trained on Wikipedia 2014 + Gigaword 5.
# khodakA: The transformation matrix (A) estimated by Khodak et al. (2018),
# trained on the same corpus as GloVe. It down-weights high-frequency
# uninformative words when computing ALC embeddings (see Rodriguez et al. 2023, p.1259).
# Both files downloaded from the conText package authors' Dropbox:
# https://www.dropbox.com/sh/jsyrag7opfo7l7i/AAB1z7tumLuKihGu2-FDmhmKa?dl=0

glove <- readRDS("glove.rds")
transform_matrix <- readRDS("khodakA.rds")

dim(glove)
dim(transform_matrix)

```

```{r}
# 3.3 Build quanteda corpus and tokens

#create the party_period covariate BEFORE building the corpus, so that conText() can find it as a docvar.
uk_clean <- uk_clean %>%
  mutate(
    # Normalise period to "Pre" / "Post" to match downstream group names
    period = dplyr::case_when(
      grepl("Pre",  period, ignore.case = TRUE) ~ "Pre",
      grepl("Post", period, ignore.case = TRUE) ~ "Post",
      TRUE ~ as.character(period)
    ),
    # Concatenate into the four group labels used everywhere downstream
    party_period = paste(party, period, sep = "_")
  )

#  sanity check
table(uk_clean$party_period)


# Build corpus — these three columns become docvars automatically
uk_corpus <- corpus(uk_clean, text_field = "text")

uk_toks <- tokens(uk_corpus, remove_punct = TRUE) %>%
  tokens_tolower()


# Confirm the covariate exists
head(docvars(uk_toks))
```

```{r}
# 3.4 Embedding regression: fair ~ party_period 
fair_model <- conText(
  formula = fair ~ party_period,
  data = uk_toks,
  pre_trained = glove,
  transform = TRUE,
  transform_matrix = transform_matrix,
  bootstrap = TRUE,
  num_bootstraps = 100,
  confidence_level = 0.95,
  window = 6L,
  case_insensitive = TRUE
)

summary(fair_model)
```

```{r}
# 3.5 Embedding regression: duty ~ party_period 
duty_model <- conText(
  formula = duty ~ party_period,
  data = uk_toks,
  pre_trained = glove,
  transform = TRUE,
  transform_matrix = transform_matrix,
  bootstrap = TRUE,
  num_bootstraps = 100,
  confidence_level = 0.95,
  window = 6L,
  case_insensitive = TRUE
)

summary(duty_model)
```

```{r}
# 3.6 Nearest neighbors: "fair" by party and period ----

# extract +/-6-word windows around every 'fair
fair_toks <- tokens_context(x = uk_toks, pattern = "fair", window = 6L)
# dfm() is document-feature matrix of those windows ; dem() isALC embedding for each window (the ALC core) ;dem_group() isaverage within each (party x period) group
fair_dfm <- dfm(fair_toks)
#It is the Khodak/Rodriguez ALC step:each window's bag-of-words vector is averaged over GloVe vectors and multiplied by the transformation matrix A
fair_dem <- dem(x = fair_dfm, pre_trained = glove, transform = TRUE, transform_matrix = transform_matrix)
#Average within each (party x period) cell to obtain 4 group ALC vectors.
fair_dem_grouped <- dem_group(x = fair_dem, groups = docvars(fair_toks, "party_period"))

#Retrieve the top-10 nearest neighbours in GloVe for each group.
fair_nns <- nns(x = fair_dem_grouped, N = 10, pre_trained = glove, as_list = TRUE)

fair_nns_df <- bind_rows(
  fair_nns$Conservative_Post,
  fair_nns$Conservative_Pre,
  fair_nns$Labour_Post,
  fair_nns$Labour_Pre
)

print(fair_nns_df, n = 40)
```


```{r}
# the nearest neighbors of "duty". the steps are similar to the "fair"'s.
duty_toks <- tokens_context(x = uk_toks, pattern = "duty", window = 6L)
duty_dfm <- dfm(duty_toks)
duty_dem <- dem(x = duty_dfm, pre_trained = glove, transform = TRUE, transform_matrix = transform_matrix)
duty_dem_grouped <- dem_group(x = duty_dem, groups = docvars(duty_toks, "party_period"))
duty_nns <- nns(x = duty_dem_grouped, N = 10, pre_trained = glove, as_list = TRUE)

duty_nns_df <- bind_rows(
  duty_nns$Conservative_Post,
  duty_nns$Conservative_Pre,
  duty_nns$Labour_Post,
  duty_nns$Labour_Pre
)

print(duty_nns_df, n = 40)
```

```{r}
# Nearest-Neighbours tables for 'duty' and 'fair'
# following Rodriguez, Spirling & Stewart (2023) Table 3 

# load package
library(gt)

#  Define a helper that turns a long NN data frame into a wide, publication-ready table 
make_nn_table <- function(nns_df, word_label) {

  # a. Within each (party × period) target, rank neighbours by cosine and keep the top 10
  nns_ranked <- nns_df %>%
    group_by(target) %>%
    arrange(desc(value), .by_group = TRUE) %>%
    slice_head(n = 10) %>%
    mutate(
      rank = row_number(),
      # combine word and cosine into one cell, e.g. "impeach (0.72)"
      cell = sprintf("%s (%.2f)", feature, value)
    ) %>%
    ungroup()
  
  # b. Flag words that appear ONLY in that column — used for bold highlighting later
  unique_flag <- nns_ranked %>%
    group_by(feature) %>%
    mutate(n_groups = n_distinct(target)) %>%
    ungroup() %>%
    mutate(is_unique = n_groups == 1) %>%
    select(target, rank, is_unique)

  # c. Pivot long → wide: rank becomes the row index, the four groups become columns
  wide <- nns_ranked %>%
    select(rank, target, cell) %>%
    pivot_wider(names_from = target, values_from = cell)

  unique_wide <- unique_flag %>%
    pivot_wider(names_from = target,
                values_from = is_unique,
                names_prefix = "uniq_")

  wide <- wide %>% left_join(unique_wide, by = "rank")
  
  # d. Enforce column order: Conservative first (Pre, Post), then Labour (Pre, Post)
  col_order <- c("rank",
                 "Conservative_Pre", "Conservative_Post",
                 "Labour_Pre",       "Labour_Post")
  col_order <- intersect(col_order, names(wide))
  uniq_cols <- paste0("uniq_", setdiff(col_order, "rank"))
  
  # e. render with gt ----
  tbl <- wide %>%
    select(all_of(col_order), all_of(uniq_cols)) %>%
    gt(rowname_col = "rank") %>%
  # title & subtitle
    tab_header(
      title = md(sprintf("**Top 10 Nearest Neighbours of '%s'**", word_label)),
      subtitle = md("Party × Period &nbsp;·&nbsp; ALC embeddings &nbsp;·&nbsp; cosine similarity in parentheses")
    ) %>%

    #  Group columns under two spanners (Conservative / Labour) for readability
    tab_spanner(label = "Conservative",
                columns = c(Conservative_Pre, Conservative_Post)) %>%
    tab_spanner(label = "Labour",
                columns = c(Labour_Pre, Labour_Post)) %>%

    # prettier column labels
    cols_label(
      Conservative_Pre  = "Pre-Brexit",
      Conservative_Post = "Post-Brexit",
      Labour_Pre        = "Pre-Brexit",
      Labour_Post       = "Post-Brexit"
    ) %>%

    #  Hide the helper flag columns — they drive styling, not display
    cols_hide(columns = all_of(uniq_cols)) %>%

    #  Bold the cells whose neighbour word is unique to that column
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_body(columns = Conservative_Pre,
                             rows = uniq_Conservative_Pre == TRUE)
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_body(columns = Conservative_Post,
                             rows = uniq_Conservative_Post == TRUE)
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_body(columns = Labour_Pre,
                             rows = uniq_Labour_Pre == TRUE)
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_body(columns = Labour_Post,
                             rows = uniq_Labour_Post == TRUE)
    ) %>%

    # f. overall typography & borders
    tab_style(
      style = cell_text(weight = "bold"),
      locations = list(cells_column_labels(), cells_column_spanners())
    ) %>%
    tab_options(
      table.font.size = 12,
      heading.title.font.size = 14,
      heading.subtitle.font.size = 10,
      column_labels.border.top.width = px(2),
      column_labels.border.bottom.width = px(1),
      table.border.top.color = "black",
      table.border.bottom.color = "black",
      row_group.border.top.color = "grey60"
    ) %>%

    # g.Footnote explaining the bold convention
    tab_source_note(
      source_note = md("*Bold* = word appears only in this (party × period) column, i.e. a semantically unique neighbour. Cosine similarity to the group ALC embedding shown in parentheses.")
    )

  tbl
}
```

```{r}
# Produce the two tables ----
duty_nn_table <- make_nn_table(duty_nns_df, "duty")
fair_nn_table <- make_nn_table(fair_nns_df, "fair")

duty_nn_table
fair_nn_table

```

#explanations of tables
#The two nearest-neighbour tables make the numerical distances concrete.

#For duty, the Conservative-Pre context is dominated by fiscal / regulatoryvocabulary, like surcharge, subsidize, excise, enact, promulgate consistent with a technical-legislative reading of "duty" as a tax or statutory obligation. In Conservative-Post, that vocabulary is largely replaced by commitment / moral_vocabulary ,like obligation, redouble, fulfil, uphold, affirm, reaffirm_. Labour-Post shares almost the entire top-ten with Conservative-Post (obligation, legislate, redouble, affirm, fulfil, reexamine, uphold, obliged, intend, reaffirm). In other words, both parties' post-Brexit "duty" shifted from fiscal duty to moral duty, and converged on the same moralised semantic neighbourhood. This is substantively important: Step 1 said Conservatives _use more_ moral language than Labour after Brexit; Step 3 adds that when both parties do use a loyalty-coded word like "duty" in the Post period, they use it in the same moralised sense. Polarisation in quantity coexists with convergence in meaning.

#For fair, the neighbour columns are noticeably noisier — species names (eucereon, chamaesphecia, pingasa) and apparent OCR artefacts (beleive, neccessary) appear in all four columns. This reflects the fact that "fair" and its true semantic neighbours (ensures, ought, obligated, presume) are high-frequency, polysemous tokens in GloVe, and much of the in-context data is dominated by formulaic parliamentary usage ("fair to say", "fair chance"). 